Intentional release of native species undermines ecological stability

### 2022-10-03 rmd/manuscript.Rmd
rmd

Intentional release, however, had little or a stabilizing influence on community dynamics when carrying capacity was sufficiently large ($K=$ `r max(list_param[[2]]$k)`; **Figures S4-S6**). In particular, intentional release increased the enhanced species with a low population growth rate ($r_1=$ `r min(list_param[[2]]$r1)`; **Figure S4**), suggesting that weak intraspecific competition within the enhanced species (= $r_1 K^{-1}$) is needed for intentional release to be effective. Our result is reasonable because some of the best evidence for successful intentional release comes from re-introduction/translocation programs of extirpated species or population enhancement of long-lived endangered species. In the Sado island (Japan), for example, the endangered Crested Ibis *Nipponia nippon* showed exponential growth of the population since the initial reintroduction of ten captive-bred individuals in 2008. Similarly, translocations of Humpback chub, a federally-listed fish species, seem highly cost-effective in enhancing their tributary populations in the Little Colorado River. Two common features exist in these successful examples. First, the receiving ecosystem remains nearly-intact (i.e., high carrying capacity). Second, the number of releases in these conservation-oriented programs is rather small (several tens to hundreds) unlike release programs for commercial and recreational species (up to several tens of millions). These conditions meet the theoretical assumptions with which intentional release led to increased abundance of the enhanced species with negligible impacts on other species. However, the preponderance of unsuccessful reintroduction programs implies that these conditions are not always sufficed. In several examples, excessive numbers of individuals are released into degraded ecosystems with absolute failures of population enhancement or establishment. Our analysis suggests that pre-release examination and restoration of environmental capacity may be a key to successful release programs with minimal impacts on other community members.

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the intentional release of masu salmon (**Figure 1E**) on the long-term stability of stream fish communities in Hokkaido, Japan. In the protected watersheds (all separated by the ocean; **Figure 1C, D**), a long-term program exists to monitor stream fish communities along with the official release records. The majority of release occurs in spring, after which salmon fry stay in freshwater for growing seasons. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` independent watersheds (see **Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of intentional release on community dynamics while accounting for potential effects of climates and local abiotic factors.

@@

Competition is a likely mechanism, as our theory assumes. Hatchery salmons are larger and more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild [@weberInteractionsHatcheryWild2003]. Indeed, field and experimental studies confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage, typically one year [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022]. Importantly, socioeconomic factors control the number of releases (e.g., budget allocation) regardless of the current condition of recipient communities. As such, released fish are probably "excessive" and may cause resource competition that would otherwise not exist.

@@

>>>>

Intentional release, however, had little or a stabilizing influence on community dynamics when carrying capacity was sufficiently large ($K=$ `r max(list_param[[2]]$k)`; **Figures S4-S6**). In particular, intentional release increased the enhanced species with a low population growth rate ($r_1=$ `r min(list_param[[2]]$r1)`; **Figure S4**), suggesting that weak intraspecific competition within the enhanced species (= $r_1 K^{-1}$) is needed for intentional release to be effective. Our result is reasonable because some of the best evidence for successful intentional release comes from re-introduction/translocation programs of extirpated species or population enhancement of long-lived endangered species. In the Sado island (Japan), for example, the endangered Crested Ibis *Nipponia nippon* showed exponential growth of the population since the initial reintroduction of ten captive-bred individuals in 2008. Similarly, translocations of Humpback chub, a federally-listed fish species, seem highly cost-effective in enhancing their tributary populations in the Little Colorado River. Two common features exist in these successful cases. First, the receiving ecosystem remains nearly-intact (i.e., high carrying capacity). Second, the number of releases in these conservation-oriented programs is rather small (several tens to hundreds) unlike release programs for commercial and recreational species (up to several tens of millions). These conditions meet the theoretical assumptions with which intentional release led to increased abundance of the enhanced species. However, the preponderance of unsuccessful reintroduction programs implies that these conditions are not always sufficed. In several examples, excessive numbers of individuals are released into degraded ecosystems with absolute failures of population enhancement or establishment. Our analysis suggests that pre-release examination and restoration of environmental capacity may be a key to successful release programs with minimal impacts on other community members.

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the intentional release of masu salmon (**Figure 1E**) on the long-term stability of stream fish communities in Hokkaido, Japan. In the protected watersheds (all separated by the ocean; **Figure 1C, D**), a long-term program exists to monitor stream fish communities along with the official release records. The release of masu salmon began in 1950s, with the annual average ranging from `r round(min(df_stock_mu$mean_stock), 2)` (no release) to `r round(max(df_stock_mu$mean_stock), 2)` million fish across watersheds during 1999 -- 2019. The majority of release occurs in spring at fish survey locations, after which salmon fry stay in freshwater for at least one year with other stream fishes (see **Figure S16** for a co-occurrence matrix in our dataset). Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` independent watersheds (see **Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of intentional release on community dynamics while accounting for potential effects of climates and local abiotic factors.

@@

To gain better insights into the driving mechanisms, we fitted a multi-species Ricker model to a subset of our time-series data at nine sites with few missing observations and no intentional release (**Methods**); thus, the parameter estimates of this model is directly comparable to those used in the theoretical Ricker model. The estimated competition coefficients $\alpha_{ij}$ ranged `r round(min(df_pd$alpha_prime), 2)` -- `r round(max(df_pd$alpha_prime), 2)` with the median of `r round(median(df_pd$alpha_prime), 2)` (Figure 3), suggesting that competition is a likely mechanism as our theory assumes. Our statistical inference is supported by previous observational studies. For example, interference competition between masu salmon and white-spotted charr is well studied. Their diets are innately overlapped (e.g., drift terrestrial insects), but they choose different prey items when they co-occur [@miyasakaFoodHabitDivergence2003]. Exploitative competition between masu salmon and benthic fish (e.g., sculpins) is also likely because subordinate masu salmon in the dominance hierarchy consumes benthic macroinvertebrates [@satoStagespecificFunctionalResponses2014], which are dominant prey items for benthic fish species. Importantly, hatchery masu salmon may intensify intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022]. Hatchery salmons are larger and more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition [@weberInteractionsHatcheryWild2003]. Indeed, field and experimental studies confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022].

@@

Importantly, socioeconomic factors control the number of releases (e.g., budget allocation) regardless of the current condition of recipient communities. As such, released fish are probably "excessive" and may cause resource competition that would otherwise not exist.
